---
title: "Correlations"
author: "P-value Data Analytics"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal:
## To assess the correlation between users' scores in January and February using Pearson correlation.

# Libraries
```{r, warning=FALSE, message=FALSE}
pacman::p_load(tidyverse, gmodels, effectsize, apaTables)
```

## From the Git Hub

1) Go to the GitHub repository link where you have the CSV file.
2) Click on the **raw** option present on the top right of the data.
3) This will open a new window in the browser.
4) The link should be like **https://raw.githubusercontent.com/..**.
5) You have to use this link to download csv file from Github.

```{r, warning=FALSE, message=FALSE}
URL <- c('https://raw.githubusercontent.com/tomushkat/R-course-for-data-management-022021/main/anovaData.csv')
dataGit <- read_csv(url(URL))
```

# Pearson correlation test

The correlation test is conducted using the **cor.test** command.
The method should be **pearson**.
```{r, warning=FALSE, message=FALSE}
Model <- cor.test(dataGit$Score[dataGit$Month == 'January'],
       dataGit$Score[dataGit$Month == 'February'],
       method = c('pearson'))
Model
```

## Effect size - R^2
The effect size is the $R^2$.
The **past0** command is for concatenating strings and numeric.
```{r, warning=FALSE, message=FALSE}
paste0(round(100 * Model$estimate ^ 2, 2), "%")
```

## Visualization

For the Visualization we first need to convert the data set tables from a long format to a short format. This is contacted using the **pivot_wider** command after we filtered only the months we want to explore and selected only the relevant variables. 

In the **ggplot** command:
1) Row 69 indicates the structure of the figure.
2) Row 70 indicates the the observations should be painted as points.
3) Row 71 indicates that there will be a line, and the **method* command indicated that the line will be linear.
4) Rows 72-75  are for aesthetics and headlines.

```{r, warning=FALSE, message=FALSE}
df_wide <- dataGit %>%
  filter(Month != 'March') %>% 
  select(ID, Month, Score) %>%
  pivot_wider(names_from = Month, values_from = Score, names_prefix = "time")


df_wide %>%
  ggplot(aes(x = timeJanuary, y = timeFebruary)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_classic() +
  ylab('Score in January') + xlab('Score in February') +
  ggtitle('Pearson correlation: Score by months Jrnuary and February') + theme(plot.title = element_text(hjust = 0.5))
```

## Correlation matrix - for several correlations at once

```{r, warning=FALSE, message=FALSE}
tableData <- dataGit %>% 
  filter(Month == 'February') %>% 
  select(monthlyGameEntry, Score, Purchase, Stage)

apa.cor.table(tableData)
```

# Spearman correlation test
For Spearman correlation change the method to 'spearman'.
## Correlation test
```{r, warning=FALSE, message=FALSE}
cor.test(dataGit$Score[dataGit$Month == 'January'],
       dataGit$Score[dataGit$Month == 'February'],
       method = c('spearman'))
```


# Chi square test
## 2 X 2

Uniting the education levels using the **mutate** and **ifelse** commands and reordering their values using the **ordered** command. 

## Data preperation
```{r, warning=FALSE, message=FALSE}
data1 <- dataGit %>% 
  filter(Month == 'March') %>% 
  mutate(Education1 = ifelse(is.na(Education), Education,
                            ifelse(Education == 'Pupil' | Education =='Soldier', 'nonAcademic', 'Academic')),
         Education1 = ordered(Education1, levels = c("nonAcademic", "Academic")))

```

### Visualization

The first row of the ggplot command paint the structure of the figure. 
The second row paint the bars and summaries the data for percentage.
The third and forth rows summaries and paint the text on top of the bars.
The fifth row paint the bars with the colors for the education.
The six row split the figure by gender.
The last row transform the Y-axis from decimal number to percent.
```{r, warning=FALSE, message=FALSE}
   data1 %>%
  ggplot(aes(x = Education1, group = Gender)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = "count") +
    geom_text(aes( label = scales::percent(..prop..),
                   y = ..prop.. ), stat= "count", vjust = -.5) +
    labs(y = "Percent", fill = "Education1") +
    facet_grid(~Gender) + theme_classic() + xlab('Education level') +
    scale_y_continuous(labels = scales::percent) 
  

```

### Chi-Square test
The chi-square calculation and table is using the **crossTable** command from the **gmodels** package.
```{r, warning=FALSE, message=FALSE}
CrossTable(data1$Education1, data1$Gender, chisq = T, format = 'SPSS')  # gmodels package
```

For assessing the effect size, first the model should be conducted using the base R function **chisq.test**, and than to use the command **effectsize**.
### Effect size - phi
```{r, warning=FALSE, message=FALSE}
Model <- chisq.test(data1$Education1, data1$Gender)

effectsize(Model, type  = 'cramers_v')
```

