---
title: "Practice"
author: "P-value Data Analytics"
date: "3/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Produce a scaled (z-score) and a non-scaled histograms of the variable 'Purchase'
 - Upload the tidyverse package
 - Load the data Data.csv from your computer and store them inside the variable 'myData'

2) Create a new data set called cleanData that that is without the outliers of the variable 'Purchase'.
Use the 'continuousOutliers' function

```{r, warning=FALSE, message=FALSE}

continuousOutliers <- function(data, lowerBound = 3, upperBpund = 3){
  
  # This function takes a vector, and replace observations that are smaller or bigger than the lower/upper bounds * standard deviations with NA
  
  Mean   <- mean(data, na.rm = TRUE)
  SD     <- sd(data, na.rm = TRUE)
  High   <- Mean + upperBound * SD # Upper Range  
  Low    <- Mean - lowerBound * SD # Lower Range
  Final  <- ifelse(data > High | data < Low, NA, data)
  return(Final)
  
}
```


5) Is there a difference between males and females in the amount of Purchase in February?
Use independent samples t-test.


6) Are there differences in the observations' scores between January and March? Examine the differences by gender separately.
Use paired samples t-tests.

